---
title: "2022 Florida Gubernatorial Election Results"
subtitle: "Interactive County-by-County Analysis and Visualization"
author: "Nicholas Tison"
date: today
format:
  html:
    theme: flatly
    toc: true
    code-fold: show
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This analysis provides an interactive visualization of county-level returns from the 2022 Florida Gubernatorial Election between **Ron DeSantis (REP)** and **Charlie Crist (DEM)**.

```{r}
#| label: setup
library(tidyverse)
library(leaflet)
library(sf)
library(maps)
library(scales)
```

## Data Loading & Preparation

We load the county election returns and calculate two-party vote shares, vote counts, and victory margins.

```{r}
#| label: load-and-clean
county_data <- read_csv("data/20221108__fl__general__county.csv")

# Filter for Governor race and compute margins by county
gov_results <- county_data |>
  filter(office == "Governor", party %in% c("DEM", "REP")) |>
  group_by(county, party) |>
  summarize(votes = sum(votes), .groups = "drop") |>
  pivot_wider(names_from = party, values_from = votes) |>
  mutate(
    total_votes = DEM + REP,
    rep_pct = REP / total_votes * 100,
    dem_pct = DEM / total_votes * 100,
    rep_margin = (REP - DEM) / total_votes * 100,
    winner = if_else(rep_margin > 0, "DeSantis (REP)", "Crist (DEM)"),
    county_clean = str_remove_all(tolower(county), "[. ]")
  )
```

## Statewide Summary

```{r}
#| label: statewide-summary
gov_results |>
  summarize(
    `DeSantis Votes` = comma(sum(REP)),
    `DeSantis Pct` = percent(sum(REP) / sum(total_votes), accuracy = 0.1),
    `Crist Votes` = comma(sum(DEM)),
    `Crist Pct` = percent(sum(DEM) / sum(total_votes), accuracy = 0.1),
    `Counties Won (DeSantis)` = sum(winner == "DeSantis (REP)"),
    `Counties Won (Crist)` = sum(winner == "Crist (DEM)")
  ) |>
  knitr::kable(align = "c")
```

## Interactive Election Map

Zoom and hover over any county to inspect county names, vote totals, candidate shares, and victory margins.

```{r}
#| label: interactive-map

# Convert Florida county boundaries to an sf object
fl_sf <- st_as_sf(maps::map("county", "florida", plot = FALSE, fill = TRUE)) |>
  st_transform(4326) |>
  mutate(
    subregion = str_remove(ID, "^florida,"),
    county_clean = str_remove_all(subregion, "[. ]")
  ) |>
  left_join(gov_results, by = "county_clean")

# Symmetrical diverging color palette (Blue = Crist, Red = DeSantis)
max_margin <- max(abs(fl_sf$rep_margin), na.rm = TRUE)
pal <- colorNumeric(
  palette = c("#2166ac", "#f7f7f7", "#b2182b"),
  domain = c(-max_margin, max_margin)
)

# Tooltip labels for hover
labels <- sprintf(
  "<strong>%s County</strong><br/>
   Winner: <b>%s</b><br/>
   DeSantis (REP): %s (%.1f%%)<br/>
   Crist (DEM): %s (%.1f%%)<br/>
   Margin: %+.1f%%",
  fl_sf$county,
  fl_sf$winner,
  comma(fl_sf$REP),
  fl_sf$rep_pct,
  comma(fl_sf$DEM),
  fl_sf$dem_pct,
  fl_sf$rep_margin
) |> lapply(htmltools::HTML)

# Build interactive Leaflet map
leaflet(fl_sf, width = "100%", height = "600px") |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    fillColor = ~pal(rep_margin),
    weight = 1.2,
    opacity = 1,
    color = "white",
    fillOpacity = 0.8,
    highlightOptions = highlightOptions(
      weight = 3,
      color = "#333",
      fillOpacity = 0.95,
      bringToFront = TRUE
    ),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "4px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) |>
  addLegend(
    pal = pal,
    values = c(-max_margin, max_margin),
    opacity = 0.8,
    title = "DeSantis Margin (%)",
    position = "bottomright"
  )
```
